---
title: "Investigating CO2 Emissions and Renewable Energy Usage"
author: "Noel Lopez, Patrick George, Riley Svensson, Ningjing Hua"
format: 
  html:
    self-contained: true
    code-tools: true
    toc: true
    number-sections: true
    code-fold: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(kableExtra)
options(scipen = 99999)
energy <- readxl::read_xlsx(here::here("renewable_energy.xlsx"))
co2 <- readxl::read_xlsx(here::here("co2.xlsx"))

```

```{r clean data}
#| output: false
convert_to_numeric <- function(x) {
  x <- str_replace_all(x, "k", "e3")
  x <- str_replace_all(x, "M", "e6")
  return(as.numeric(x))
}


co2_clean <- co2 |>
  select(country, `1989`:`2017`) |>
  mutate(across(.cols = `1989`:`2017`, 
                ~ convert_to_numeric(.))) |>
  pivot_longer(cols = !country,
               names_to = "year",
               values_to = "Total_CO2_Emissions")

energy_clean <- energy |>
  pivot_longer(cols = !country,
               names_to = "year",
               values_to = "Percent_Consumption_Renewable")

master <- inner_join(co2_clean, energy_clean) |>
  drop_na()

master |>
  distinct(year) |>
  count()

master |>
  distinct(country) |>
  count()
```

# Introduction

Climate change is a global challenge jeopardizing humanity's future. This purpose of this project is to investigate the relationship between CO2 emissions, one of the largest drivers of climate change creating greenhouse gasses, and the use of renewable energy such as solar, wind, hydroelectric or nuclear energy. The data to be used in this project incorporates data from two different sources, one measuring total carbon dioxide emissions per country over the years and one measuring proportion of energy use which is renewable over the years. `Total_CO2_Emission` is in 1000 tons of CO2 and `Percent_Consumption_Renewables` is a percentage of total energy use produced by renewable sources per country over the years. There are 207 distinct countries observed over 29 years. The relationship between Total_CO2_Emission and Percent_Consumption_Renewables is hypothesized to be negative whereas a country uses more renewable energy they produce less CO2. Through this investigation hopefully more information can be learned about climate change and potential solutions to this global problem affecting every living and non-living object.

# Materials and Methods

As mentioned in the introduction, the master dataset to be used for analysis in this project incorporates data from two sets, one measuring the total carbon dioxide emissions, per country over the years and the other measuring the proportion of energy use, which is classified as renewable energy over the same time. The data was sourced from an online database titled Gapminder which has been a reliable provider of data since 2005, in hopes of promoting global sustainability through easily accessible information. Before analysis, the data had to be properly cleaned and wrangled.

The CO2 dataset contained values in the form of 25.3k and 4.9M, signaling the units of thousands and millions of units. The first step in the data cleaning process was to replace these letters with proper units to create a numeric variable, which enabled the performance of calculations on the data - a function titled was created for this purpose. The data was then pivoted to be in tidy form and the years of interest (1989 -- 2017) were selected. Once pivoted, the data sets were joined by year and country to produce a master data set which shall be used for model fitting, plotting, and subsequent analysis. A final decision was made to drop all the Na's in the master set due to a few reasons. The missing values typically occurred in the `Percent_Consumption_Renewables` variable across the earlier years and usually in small countries. Whether these Na's were include due to lack of observation or for simply not having any renewable energy production or consumption it is unclear. Because of a significant sample size after dropping Na's, and because of this knowledge gap and lack of patterns about the data, dropping Na's is appropriate and other data imputation methods such as cell-mean imputation would not be a proper solution as this introduces bias and would make a time-series graph incorrect.

All the data cleaning, linear regression models, and subsequent analysis was conducted using R code.

```{r}
datatable((head(master, n = 50)),
          caption = 'Interactive Preview of Data Set')
```

# Analysis and Plots

```{r}
#| output: false
# linear regression analysis on Total_CO2_Emissions and Percent_Consumption_Renewable
energy_emissions_model <- lm(`Total_CO2_Emissions` ~ `Percent_Consumption_Renewable`,
                             data = master)
summary(energy_emissions_model)
tidy(energy_emissions_model) 
```

Simple regression equation:

$$ \hat{y} = 217719 - 2590 * Percent\,Consumption\,Renewable $$

Based on the linear regression model investigating the relationship between total carbon dioxide emissions and proportion of renewable energy, we can see that there is a negative relationship. This means that generally, the higher the proportion of renewable energy a country uses, the lower the total carbon dioxide emissions. The response variable in this equation is the country's Total CO2 emissions in 1000 of tons while the explanatory variable is the percentage of energy a country uses which is classified as renewable. To interpret the coefficient of -2590 this means that for each additional 1% of renewable energy consumption used a country's predicted total CO2 emission in thousands of tons decreases by 2590 on average.

The method used to predict this model was simple linear regression. This method involves predicting the values of one variable, based on another, through producing a straight line minimizing the value for the sum of squared residuals. Based on the linear regression model investigating the relationship between total carbon dioxide emissions and proportion of renewable energy, we can see that there is a negative relationship. This means that generally, the higher the proportion of renewable energy a country uses, the lower the total carbon dioxide emissions. The response variable in this equation is the country's Total CO2 emissions in 1000 of tons while the explanatory variable is the percentage of energy a country uses which is classified as renewable. To interpret the coefficient of -2590 this means that for each additional 1% of renewable energy consumption used a country's predicted total CO2 emission in thousands of tons decreases by 2590 on average.

```{r}
#| fig-align: center
# Plot 1
master |> 
ggplot(aes(x = Percent_Consumption_Renewable, 
           y = Total_CO2_Emissions)
       ) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme(legend.position = "none") +
  labs(x = "Renewable energy consumption percentage", 
       y = "", 
       title = "Relationships between Renewable Energy Consumption and CO2 Emissions among different contries", 
       subtitle = "Total CO2 Emissions/1000 tonnes") 
```

```{r}
#| fig-align: center
# Plot 1
master |> 
ggplot(aes(x = Percent_Consumption_Renewable, 
           y = Total_CO2_Emissions, color=country)
       ) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme(legend.position = "none") +
  labs(x = "Renewable energy consumption percentage", 
       y = "", 
       title = "Relationships between Renewable Energy Consumption and CO2 Emissions among different contries", 
       subtitle = "Total CO2 Emissions/1000 tonnes") +
  annotate("text", x = 25, y = 6200000, label = "China", size=2) +
  annotate("text", x = 8, y = 6500000, label = "US",size=2) +
  annotate("text", x = 43, y = 2200000, label = "India", size=2) +
  annotate("text", x = 3, y = 2900000, label = "Russia", size=2) + ylim(900000, 10000000)
```

```{r}
master |>
  group_by(year) |>
  summarize(avg_co2 = mean(Total_CO2_Emissions)) |>
  ggplot(aes(x = year , y = avg_co2)) +
  geom_point() 

master |>
  group_by(year) |>
  summarize(avg_percent = mean(Percent_Consumption_Renewable)) |>
  ggplot(aes(x = year, y = avg_percent)) +
  geom_point()
```

```{r}
energy_emissions_model |>
  augment() |>
  summarize(`Variance of Fitted` = var(.fitted),
            `Variance of Residuals` = var(.resid),
            `Variance of Total_CO2_Emissions` = var(Total_CO2_Emissions)) |>
  kable(caption = 'Model Fit') |>
  kable_styling(bootstrap_options = c('striped', 'bordered'))

```

The proportion of variability in the response values that was accounted for by the model was very small at about 1.573 percent. This suggests a poor model, where very little of the variation in the response is truly explained by the exlpanatory variable. This suggests that for all countries around the world there are many more factors influecing the amount of CO2 emissions produced rather than the proportion of renewable energy used. When investigating individual countries, however, for example China or India, we see high R squared suggesting that more developed contries follow a better trend and renewable energy usage has a bigger impact.

# Discussion of Results

# References

https://www.gapminder.org/data/

# For reference

```{r}
#| output: false
## This is just for model using Chinese data
china <- master |>
  filter(country == 'China')

china_model <- lm(`Total_CO2_Emissions` ~ `Percent_Consumption_Renewable`,
                             data = china)
summary(china_model)
tidy(china_model) 

```

```{r}
energy_emissions_model |> 
  augment() |> 
ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() 
```
