---
title: "Investigating CO2 Emissions and Renewable Energy Usage"
author: "Noel Lopez, Patrick George, Riley Svensson, Ningjing Hua"
format: 
  html:
    self-contained: true
    code-tools: true
    toc: true
    number-sections: true
    code-fold: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(kableExtra)
library(gridExtra)
library(patchwork)
options(scipen = 99999)
energy <- readxl::read_xlsx(here::here("renewable_energy.xlsx"))
co2 <- readxl::read_xlsx(here::here("co2.xlsx"))

```

```{r clean data}
#| output: false
convert_to_numeric <- function(x) {
  x <- str_replace_all(x, "k", "e3")
  x <- str_replace_all(x, "M", "e6")
  return(as.numeric(x))
}


co2_clean <- co2 |>
  select(country, `1989`:`2017`) |>
  mutate(across(.cols = `1989`:`2017`, 
                ~ convert_to_numeric(.))) |>
  pivot_longer(cols = !country,
               names_to = "year",
               values_to = "Total_CO2_Emissions")

energy_clean <- energy |>
  pivot_longer(cols = !country,
               names_to = "year",
               values_to = "Percent_Consumption_Renewable")

joined <- inner_join(co2_clean, energy_clean) |>
  drop_na()

joined |>
  distinct(year) |>
  count()

joined |>
  distinct(country) |>
  count()

master <- joined|>
  rename(Year = year) |>
  group_by(Year) |>
  summarize(`Average Emissions` = mean(`Total_CO2_Emissions`),
            `Average Proportion Renewawable`
            = mean(`Percent_Consumption_Renewable`))
```

# Introduction

Climate change is a global challenge jeopardizing humanity's future. This purpose of this project is to investigate the relationship between CO2 emissions, one of the largest drivers of climate change creating greenhouse gasses, and the use of renewable energy such as solar, wind, hydroelectric or nuclear energy. The data to be used in this project incorporates data from two different sources, one measuring total carbon dioxide emissions per country over the years and one measuring proportion of energy use which is renewable over the years. `Total_CO2_Emission` is in 1000 tons of CO2 and `Percent_Consumption_Renewables` is a percentage of total energy use produced by renewable sources per country over the years. There are 207 distinct countries observed over 29 years. The relationship between Total_CO2_Emission and Percent_Consumption_Renewables is hypothesized to be negative whereas a country uses more renewable energy they produce less CO2. Through this investigation hopefully more information can be learned about climate change and potential solutions to this global problem affecting every living and non-living object.

# Materials and Methods

As mentioned in the introduction, the master dataset to be used for analysis in this project incorporates data from two sets, one measuring the total carbon dioxide emissions, per country over the years and the other measuring the proportion of energy use, which is classified as renewable energy over the same time. The data was sourced from an online database titled Gapminder which has been a reliable provider of data since 2005, in hopes of promoting global sustainability through easily accessible information. Before analysis, the data had to be properly cleaned and wrangled.

The CO2 dataset contained values in the form of 25.3k and 4.9M, signaling the units of thousands and millions of units. The first step in the data cleaning process was to replace these letters with proper units to create a numeric variable, which enabled the performance of calculations on the data - a function titled was created for this purpose. The data was then pivoted to be in tidy form and the years of interest (1989 -- 2017) were selected. Once pivoted, the data sets were joined by year and country to produce a master data set which shall be used for model fitting, plotting, and subsequent analysis. A final decision was made to drop all the Na's in the master set due to a few reasons. The missing values typically occurred in the `Percent_Consumption_Renewables` variable across the earlier years and usually in small countries. Whether these Na's were include due to lack of observation or for simply not having any renewable energy production or consumption it is unclear. Because of a significant sample size after dropping Na's, and because of this knowledge gap and lack of patterns about the data, dropping Na's is appropriate and other data imputation methods such as cell-mean imputation would not be a proper solution as this introduces bias and would make a time-series graph incorrect.

The method used to predict this model was simple linear regression. This method involves predicting the values of one variable, based on another, through producing a straight line minimizing the value for the sum of squared residuals. All the data cleaning, linear regression models, and subsequent analysis was conducted using R code.

```{r}
datatable((head(master, n = 50)),
          caption = 'Interactive Preview of Data Set')
```

# Analysis and Discussion

```{r}
#| output: false
energy_emissions_model <- lm(`Average Emissions` ~ 
                             `Average Proportion Renewawable`,
                             data = master)
summary(energy_emissions_model)
tidy(energy_emissions_model)

```

Simple regression equation:

$$ \hat{y} = 693309.6 - 17770.2 * Average\,Percent\,Renewable $$

Above is the linear regression equation based on the model predicting the response variable, Average Carbon Dioxide Emission ($\hat{y}$), based off the explanatory variable, Average Proportion of Renewable Energy consumption. The coefficient on `Average Percent Renewable` is extremely negative sitting at -17770.2 meaning that for each one percent increase in the `Average Percent Renewable` energy used the average amount of CO2 emissions in thousands of tons decreases by 17770.2. The slope coefficient of 693309.6 means that when the `Average Percent Renewable` is zero such that there is no renewable enerygy being used at all on average, the predicted average CO2 emissions would be 693309.6 thousand tons. 

```{r}
#| fig-align: center
# Plot 1
raw_graph <- master |> 
ggplot(aes(x = `Average Proportion Renewawable`, 
           y = `Average Emissions`)
       ) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme(legend.position = "none") +
  labs(x = " Average Renewable energy consumption percentage", 
       y = "", 
       title = "Relationships between Average Renewable Energy Consumption and \nAverage CO2 Emissions among different contries", 
       subtitle = "Average CO2 Emissions/1000 tonnes") +
  theme(plot.title.position = "plot")

raw_graph

```

The graph displayed on the top demonstrates the relationship between `Total_CO2_Emission` and `Percent_Renewable_Energy`, shown by plotting the raw data. The distribution illustrates a negative pattern with considerable noise, explained by the smallest countries and relative values of Total Co2 Emissions. To better visualize how the relationship between Co2 Emissions and `Percent_Renewable_Energy` differentiates for the four largest producers of Co2, we filtered to only include India, Russia, US, and China in the second plot. In doing this, we were able to take out countries with insignificant Co2 emissions, as the model better explained the variability ($R^{2}$) in Co2 emissions for the top countries. Essentially, we were able to see that there needs to be a minimum amount of `Total_CO2_Emission` data, in order for the model to be accurate in making conclusions about the variable's relationship.

```{r}
#| fig-align: center

co2_by_year_graph <- master |>
  ggplot(aes(x = Year , y = `Average Emissions`)) +
  geom_point() +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  labs(x = "Year", 
       y = "", 
       title = "Average CO2 Emissions from 1989-2017", 
       subtitle = "Average CO2 Emissions/1000 tonnes")+
  theme(plot.title.position = "plot")

energy_by_year_graph<- master |>
  ggplot(aes(x = Year, y = `Average Proportion Renewawable`)) +
  geom_point() +
   scale_x_discrete(guide = guide_axis(n.dodge=2))+
    labs(x = "Year", 
       y = "", 
       title = "Average Percentage of Renewable Energy from 1989-2017",
       subtitle = "Average Percentage of Renewable Energy")+
      theme(plot.title.position = "plot")

grid.arrange(co2_by_year_graph, energy_by_year_graph)
```

As shown by the two distributions of `Total_CO2_Emission` and Percent Renewable Energy over time, the relationship follows the negative relationship as expected. This means that as the predictor variable, Percent of Renewable Energy decreases throughout the years, the response variable, `Total_CO2_Emission` increases in turn. It's worth noting that during the timespan of 1989 to 2017, the total CO2 emissions has been increasing as expected, and is likely due to the extreme population growth over this time. On the other hand, the Percent of Renewable Energy has shockingly been declining in modern years, which is likely due to the varying definition of renewable energy, it only encompasses solar, wind, geothermal, and hydro powered. Before this discovery, we had assumed that energy sources such as nuclear power and natural gas, fell under the umbrella of renewable energy, which was ultimately false. Following this analysis, we were able to understand that the value of CO2 emissions was expanding, due to population increase, while the amount of renewable energy has actually been declining, as this only included specific sources of energy which have lost popularity today.

```{r}
energy_emissions_model|>
  augment() |>
  summarize(`Variance of Fitted` = var(.fitted),
            `Variance of Residuals` = var(.resid),
            `Variance of Average CO2 Emissions` = var(`Average Emissions`)) |>
  kable(caption = 'Model Fit',
        digits = 3,
        format.args = list(big.mark = ",")) |>
  kable_styling(bootstrap_options = c('striped', 'bordered'))

```

```{r}
energy_emissions_model |> 
  augment() |> 
  ggplot(aes(x=.fitted, y = .resid))+
  geom_point()
```

The proportion of variability in the response values that was accounted for by the model was very small at about 1.573 percent. This suggests a poor model, where very little of the variation in the response is truly explained by the explanatory variable. This suggests that for all countries around the world there are many more factors influencing the amount of CO2 emissions produced rather than the proportion of renewable energy used. When investigating individual countries, however, for example China or India, we see high $R^{2}$ suggesting that more developed countries follow a better trend and renewable energy usage has a bigger impact.

# References

https://www.gapminder.org/data/


```{r}
noise <- function(x, mean = 0, sd){
  x + rnorm(length(x), 
            mean, 
            sd)
}
```

```{r}
master_predict <- predict(energy_emissions_model)
master_sigma <- sigma(energy_emissions_model)

sim_response <- tibble(sim_emissions = noise(master_predict, 
                                          sd = master_sigma))






raw_graph <- master |> 
ggplot(aes(x = `Average Proportion Renewawable`, 
           y = `Average Emissions`)
       ) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme(legend.position = "none") +
  labs(x = " Avg Renewable energy consumption percentage", 
       y = "", 
       title = "Relationships between Avg Renewable Energy Consumption and \nAvg CO2 Emissions among different contries", 
       subtitle = "Avg CO2 Emissions/1000 tonnes") +
  theme(plot.title.position = "plot")




```

```{r}
obs_emissions <- master |>
  ggplot(aes(x = `Average Emissions`)) +
  geom_histogram(binwidth = 3000) +
  labs(x = "Observed Emissions",
       y = "",
       subtitle = "Count") +
  theme_bw()


sim_emissions_graph <- sim_response |>
  ggplot(aes(x = sim_emissions)) +
  geom_histogram(binwidth = 3500) +
  labs(x = "Simulated Emissions",
       y = "",
       subtitle = "Count") +
  theme_bw()

obs_emissions + sim_emissions_graph

```

```{r}
sim_data <- master |> 
  filter(!is.na(`Average Emissions`), 
         !is.na(`Average Proportion Renewawable`)
         ) |> 
  select(`Average Emissions`, `Average Proportion Renewawable`) |> 
  bind_cols(sim_response)


raw_graph <- master |> 
ggplot(aes(x = `Average Proportion Renewawable`, 
           y = `Average Emissions`)
       ) +
  geom_point() + 

  theme(legend.position = "none") +
  labs(x = " Avg Renewable energy consumption percentage", 
       y = "", 
       title = "Observed Relationships between Avg Renewable Energy Consumption and \nAvg CO2 Emissions among different contries", 
       subtitle = "Avg CO2 Emissions/1000 tonnes") +
  theme(plot.title.position = "plot")

sim_master_graph <- sim_data |> 
ggplot(aes(x = `Average Proportion Renewawable`, 
           y = sim_emissions)
       ) +
  geom_point() + 
  theme(legend.position = "none") +
  labs(x = " Avg Renewable energy consumption percentage", 
       y = "", 
       title = "Simulated Relationships between Avg Renewable Energy Consumption and \nAvg CO2 Emissions among different contries", 
       subtitle = "Simulated Avg CO2 Emissions/1000 tonnes") +
  theme(plot.title.position = "plot")

raw_graph + sim_master_graph
```

```{r}
# Check the similarity between the simulated data and observed data
sim_data |> 
  ggplot(aes(x = sim_emissions, 
             y = `Average Emissions`)
         ) + 
  geom_point() + 
   labs(x = "Simulated Avg CO2 Emissions/1000 tonnes", 
        y = "",
        subtitle = "Avg CO2 Emissions/1000 tonnes" ) + 
  geom_abline(slope = 1,
              intercept = 0, 
              color = "steelblue",
              linetype = "dashed",
              lwd = 1.5) +
  theme_bw()
```

```{r}
# check how the simulated data fit the observed data, especially the R square.
lm(`Average Emissions` ~ sim_emissions, 
   data = sim_data
   ) |> 
  glance() 
```

```{r}
# get the R squared for the simulated data fit for observed data.
sim_r2 <- lm(`Average Emissions` ~ sim_emissions, 
             data = sim_data
             ) |> 
  glance() |> 
  select(r.squared) |> 
  pull()
sim_r2
```

```{r}
# Created 1000 simulated dataset
nsims <- 1000
sims <- map_dfc(.x = 1:nsims,
                .f = ~ tibble(sim = noise(master_predict, 
                                          sd = master_sigma)
                              )
                )

head(sims)
```

```{r}
# clean the colnames
colnames(sims) <- colnames(sims) |> 
  str_replace(pattern = "\\.\\.\\.",
                  replace = "_")
head(sims)
```

```{r}
# bind 1000 simulated dataset and the observed dataset together
sims <- master |> 
  filter(!is.na(`Average Emissions`), 
         !is.na(`Average Proportion Renewawable`)) |> 
  select(`Average Emissions`) |> 
  bind_cols(sims)

head(sims)
```

```{r}
# mapping to get 1000 simulated R squared.
sim_r_sq <- sims |> 
  map(~ lm(`Average Emissions` ~ .x, data = sims)) |> 
  map(glance) |> 
  map_dbl(~ .x$r.squared)

head(sim_r_sq)
```

```{r}
# to see the distribution of the 1000 simulated R square.
tibble(sims = sim_r_sq) |> 
  ggplot(aes(x = sims)) + 
  geom_histogram(binwidth = 0.025) +
  labs(x = expression("Simulated"~ R^2),
       y = "",
       subtitle = "Number of Simulated Models")
#The distribution of these values will tell if our assumed model does a good job of producing data similar to what was observed. If the model produces data similar to what was observed, we would expect values near 1.
# Our model is R square is around 0.8.
```
